### 2022-06-01 Scripts/Evaluations/identifying_viral_contigs_20220521.Rmd
Added updated rule set.

---
title: "Viral Sequence Sorting Evaluation"
output: html_notebook
---

```{r}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
```

# All Viral Contigs

## checkV
```{r}
checkV <- read_tsv("../TestingSetOutput/quality_summary_20220521.tsv")
checkV <- separate(checkV, col = contig_id, into = c("seqtype", "contig"), sep="--")
checkV$contig <- sub("\\.", "_", checkV$contig)
```

## VIBRANT
```{r}
vb_c <- read_tsv("../TestingSetOutput/VIBRANT_genome_quality_metagenomic_testing_set_20220521.tsv")
vb_c <- separate(vb_c, col = scaffold, into = c("seqtype", "contig"), sep="--")
vb_c$contig <- sub("\\.", "_", vb_c$contig)
vb_c$contig <- gsub("_fragment_.*", "", vb_c$contig)
vb_c <- separate(vb_c, col=contig, into="contig", remove=T, sep = " ")
vb_c <- vb_c[!duplicated(vb_c$contig),]
```

## DeepVirFinder
```{r}
dvf_c <- read_tsv("../TestingSetOutput/dvf_output_20220521.tsv")
dvf_c <- separate(dvf_c, col = name, into = c("seqtype", "contig"), sep="--")
dvf_c$contig <- sub("\\.", "_", dvf_c$contig)
dvf_c$bh_pvalue <- p.adjust(dvf_c$pvalue, method="BH")
```

## VirSorter2 
```{r}
vs2_c <- read_tsv("../TestingSetOutput/final-viral-score_20220521.tsv")
vs2_c <- separate(vs2_c, col = seqname, into = c("seqtype", "contig"), sep="--")
vs2_c <- separate(vs2_c, col = contig, into = c("contig", "vs2type"), "\\|\\|")
vs2_c$contig <- sub("\\.", "_", vs2_c$contig)
vs2_c <- vs2_c[!duplicated(vs2_c$contig),]
```

## Kaiju
(parsing when reading the file in "fails" for sequences that weren't classified)
```{r}
kj_c <- read_tsv("../VSTE/ToolOutput/metagenomic_testing_set_nrandeuk.kaiju.names.out", col_names = F)
colnames(kj_c) <- c("Classified", "Contig", "NCBI_taxon", "len", "ID_best", "IDs_all", "Seq", "Name")
kj_c <- separate(kj_c, col = Contig, into = c("seqtype", "contig"), sep="--")
kj_c <- separate(kj_c, col = Name, into = c("Kaiju_Viral","Kingdom"), sep=";")
kj_c$contig <- sub("\\.", "_", kj_c$contig)
```

```{r}
table(kj_c$seqtype, kj_c$Kaiju_Viral)
```

## VirSorter
```{r}
vs_c <- read_csv("../TestingSetOutput/VIRSorter_global-phage-signal.csv")
vs_c$`## Contig_id` <- sub("VIRSorter_", "", vs_c$`## Contig_id`)
vs_c <- separate(vs_c, col = `## Contig_id`, into = c("seqtype", "contig"), sep="--")
vs_c$contig <- sub("-circular", "", vs_c$contig)
vs_c <- vs_c[!duplicated(vs_c$contig),]
```

## Merging

```{r}
viruses <- full_join(x=vs2_c, y=dvf_c, by = c("contig", "seqtype"))
viruses <- full_join(x=viruses, y=checkV, by = c("contig", "seqtype"))
viruses <- full_join(x=viruses, y=vb_c, by = c("contig", "seqtype"))
viruses <- full_join(x=viruses, y=vs_c, by = c("contig", "seqtype"))
viruses <- full_join(x=viruses, y=kj_c, by = c("contig", "seqtype"))
```

calculate percent viral/host/unknown
```{r}
viruses$percent_host <- viruses$host_genes/viruses$gene_count*100
viruses$percent_viral <- viruses$viral_genes/viruses$gene_count*100
viruses$percent_unknown <- 100-(viruses$host_genes+viruses$viral_genes)/viruses$gene_count*100
```

get rid of NAs for downstream processing
```{r}
viruses$Quality[is.na(viruses$Quality)] <- 0
viruses$score[is.na(viruses$score)] <- 0
viruses$bh_pvalue[is.na(viruses$bh_pvalue)] <- 0
viruses$viral[is.na(viruses$viral)] <- 0
viruses$hallmark[is.na(viruses$hallmark)] <- 0
viruses$percent_viral[is.na(viruses$percent_viral)] <- 0
viruses$percent_unknown[is.na(viruses$percent_unknown)] <- 0
viruses$Kaiju_Viral[is.na(viruses$Kaiju_Viral)] <- "unknown"
viruses$Kingdom[is.na(viruses$Kingdom)] <- "unknown"
viruses$Category[is.na(viruses$Category)] <- 0
```

##################################################

```{r}
getting_viral_set_1 <- function(input_seqs, include_vibrant=FALSE, include_virsorter2=FALSE,
                              include_deepvirfinder=FALSE, include_checkV=FALSE,
                              include_kaiju=FALSE, include_virsorter=FALSE) {

  keep_score <- rep(0, nrow(input_seqs))

  if (include_vibrant) {
    keep_score[input_seqs$Quality=="high quality draft"] <- keep_score[input_seqs$Quality=="high quality draft"] + 1
    keep_score[input_seqs$Quality=="medium quality draft"] <- keep_score[input_seqs$Quality=="medium quality draft"] + 1
    keep_score[input_seqs$Quality=="low quality draft" & input_seqs$provirus=="Yes"] <- keep_score[input_seqs$Quality=="low quality draft" & input_seqs$provirus=="Yes"] + 0.5
  }

  if (include_virsorter2) {
    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.5
    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
  }

  if (include_virsorter) {
    keep_score[input_seqs$Category==1] <- keep_score[input_seqs$Category==1] + 1
    keep_score[input_seqs$Category==2] <- keep_score[input_seqs$Category==2] + 1
    keep_score[input_seqs$Category==3] <- keep_score[input_seqs$Category==3] + 0.5
    keep_score[input_seqs$Category==4] <- keep_score[input_seqs$Category==4] + 1
    keep_score[input_seqs$Category==5] <- keep_score[input_seqs$Category==5] + 1
    keep_score[input_seqs$Category==6] <- keep_score[input_seqs$Category==6] + 0.5
  }

  if (include_deepvirfinder) {
    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
#   keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 0.5
  }

  if (include_kaiju) {
    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 1
    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 1
  }

  if (include_checkV) {
    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0
    keep_score[input_seqs$contig_length>50000 & input_seqs$hallmark==0] <- 0
    keep_score[input_seqs$percent_unknown>=75] <- keep_score[input_seqs$percent_unknown>=75] + 0.5
  }

  return(keep_score)

}

```

```{r}
assess_performance <- function(seqtype, keep_score) {

  truepositive <- rep("not viral", length(seqtype))
  truepositive[seqtype=="virus"] <- "viral"

  #make confusion matrix
  confusion_matrix <- rep("true negative", length(keep_score))
  confusion_matrix[truepositive=="viral" & keep_score<=1] <- "false negative"
  confusion_matrix[truepositive=="viral" & keep_score>=1] <- "true positive"
  confusion_matrix[truepositive=="not viral" & keep_score>=1] <- "false positive"

  precision <- table(confusion_matrix)[4]/(table(confusion_matrix)[4]+table(confusion_matrix)[2])

  recall <- table(confusion_matrix)[4]/(table(confusion_matrix)[4]+table(confusion_matrix)[1])

  F1 <- 2*precision*recall/(precision+recall)

  performance <- c(precision, recall, F1)
  names(performance) <- c("precision", "recall", "F1")

  return(performance)
}
```

```{r}
comp_pipelines <- tibble(pipeline="example", precision=0, recall=0, F1=0)
```

## just vibrant
```{r}
viruses$keep_score_vb <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)

assess_performance(viruses$seqtype, viruses$keep_score_vb)
```

## vibrant + checkV
```{r}
viruses$keep_score_vb_cv <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)

assess_performance(viruses$seqtype, viruses$keep_score_vb_cv)
```

## vibrant + checkV + kaiju
```{r}
viruses$keep_score_vb_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)

assess_performance(viruses$seqtype, viruses$keep_score_vb_cv_kj)
```

## vibrant + checkV + kaiju + deepvirfinder
```{r}
viruses$keep_score_vb_cv_kj_dvf <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)

assess_performance(viruses$seqtype, viruses$keep_score_vb_cv_kj_dvf)
```

## vibrant + virsorter2 + checkV + kaiju
```{r}
viruses$keep_score_vb_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)

assess_performance(viruses$seqtype, viruses$keep_score_vb_vs2_cv_kj)
```

#all: vibrant + virsorter2 + checkV + kaiju + deepvirfinder
```{r}
viruses$keep_score_all <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)

assess_performance(viruses$seqtype, viruses$keep_score_all)
```

## just virsorter2
```{r}
viruses$keep_score_vs2 <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)

assess_performance(viruses$seqtype, viruses$keep_score_vs2)
```

## virsorter2 + checkV
```{r}
viruses$keep_score_vs2_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)

assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv)
```

## virsorter2 + checkV + kaiju
```{r}
viruses$keep_score_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)

assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv_kj)
```

## virsorter2 + checkV + kaiju + deepvirfinder
```{r}
viruses$keep_score_vs2_cv_kj_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)

assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv_kj_dvf)
```

## just checkV
```{r}
viruses$keep_score_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)

assess_performance(viruses$seqtype, viruses$keep_score_cv)
```

## checkV + kaiju
```{r}
viruses$keep_score_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)

assess_performance(viruses$seqtype, viruses$keep_score_cv_kj)
```

## just kaiju
```{r}
viruses$keep_score_k <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = T,
                                            include_deepvirfinder = F,
                                            include_virsorter = F)

assess_performance(viruses$seqtype, viruses$keep_score_k)
```

## just deepvirfinder
```{r}
viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = T)

assess_performance(viruses$seqtype, viruses$keep_score_dvf)
```

## deepvirfinder + checkV
```{r}
viruses$keep_score_dvf_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = T)

assess_performance(viruses$seqtype, viruses$keep_score_dvf_cv)
```

## deepvirfinder + checkV + kaiju
```{r}
viruses$keep_score_dvf_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T,
                                            include_virsorter = F)

assess_performance(viruses$seqtype, viruses$keep_score_dvf_cv_kj)
```

## virsorter2 + checkV + kaiju
```{r}
viruses$keep_score_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)

assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv_kj)
```

## just virsorter
```{r}
viruses$keep_score_vs <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)

assess_performance(viruses$seqtype, viruses$keep_score_vs)
```

## virsorter + checkV
```{r}
viruses$keep_score_vs_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)

assess_performance(viruses$seqtype, viruses$keep_score_vs_cv)
```

## virsorter + checkV + kaiju
```{r}
viruses$keep_score_vs_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)

assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj)
```

## virsorter + checkV + kaiju + vibrant + deepvirfinder + virsorter2
```{r}
viruses$keep_score_vs_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T,
                                            include_virsorter = T)

assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj)
```

## virsorter + checkV + kaiju + vibrant
```{r}
viruses$keep_score_vs_cv_kj_vb <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)

assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj_vb)
```

## virsorter + checkV + kaiju + virsorter2 + vibrant
```{r}
viruses$keep_score_vs_cv_kj_vb_vs2 <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)

assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj_vb_vs2)
```

## virsorter + checkV + kaiju + virsorter2
```{r}
viruses$keep_score_vs_cv_kj_vs2_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T,
                                            include_virsorter = T)

assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj_vs2_dvf)
```

# Plot of how many methods identify a contig
```{r}
viruses_high <- viruses[viruses$keep_score_all>=1,]
viruses_high_mod <- viruses_high %>% select(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj)
viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
viruses_high_mod <- as_tibble(viruses_high_mod)

sm <- as.data.frame(rbind(c(0,table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vb_cv>0,]))),
                          c(table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_dvf_cv>0,]))),
                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vs2_cv>0,])),
                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_cv_kj>0,]))
                          )
)

sm$method <- c("vibrant", "deepvirfinder", "virsorter2", "kaiju")

colnames(sm) <- c("one","two","three","all","method")

sm
```

```{r}
sm_m <- melt(sm)
colnames(sm_m) <- c("primary_method", "number_methods", "number")
```

```{r}
ggplot(sm_m, aes(x=primary_method, y=number,
                   fill=number_methods)) +
  geom_bar(stat="identity") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_manual(name = 'Number of Methods',
                     values = alpha(c(viridis(5)), 1)) +
  xlab("Primary Method") +
  ylab("Count of Viral Contigs")
```

# Visualizing which tools led to a contig being selected
(alluvial plot)
with help from: https://www.bioinfo-scrounger.com/archives/alluvial-and-Sankey/

```{r}
alluvial_data <- viruses %>% select(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj) %>% 
  apply(c(1,2), function(x) {if (x >= 1) {x <- 1} else if (x==0.5) {x <- 0.5} else {x <- 0}}) %>%
  as_tibble() %>%
  count(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj)
```

```{r}
alluvial_data$cum_sum <- alluvial_data %>% select(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj) %>% 
  rowSums()
```

```{r}
alluvial_data$viral <- "not viral"
alluvial_data$viral[alluvial_data$cum_sum>=1] <- "one method"
alluvial_data$viral[alluvial_data$cum_sum>=2] <- "two methods"
alluvial_data$viral[alluvial_data$cum_sum>=3] <- "three methods"
alluvial_data$viral[alluvial_data$cum_sum>=4] <- "all methods"

alluvial_data$viral <- factor(alluvial_data$viral, levels=c("not viral", "one method",
                                                            "two methods", "three methods",
                                                            "all methods"))
```

```{r}
alluvial_data <- alluvial_data %>% 
  mutate(across(c(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj), as.factor))
```

```{r}
head(alluvial_data)
```

```{r}
library(ggalluvial)
```

```{r}
ggplot(data = alluvial_data,
       aes(axis1 = as.character(keep_score_vb_cv), 
           axis2 = as.character(keep_score_vs2_cv),
           axis3 = as.character(keep_score_dvf_cv),
           axis4 = as.character(keep_score_cv_kj),
           y = n)) +
  geom_alluvium(aes(fill = viral)) +
  geom_stratum() +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum))) +
  scale_fill_viridis_d() + scale_x_discrete(limits = c("VIBRANT", "VirSorter2", "DeepVirFinder", "Kaiju"),
                                  expand = c(0.05, 0.05, 0.05, 0.05)) +
  ylab("Number of Sequences") +
  theme_classic()
```

# Ignoring True Negatives and visualizing flows
(alluvial plot)
with help from: https://www.bioinfo-scrounger.com/archives/alluvial-and-Sankey/

```{r}
viruses$confusion_matrix <- "true negative"
viruses$confusion_matrix[viruses$seqtype=="virus" & viruses$keep_score_all<=1] <- "false negative"
viruses$confusion_matrix[viruses$seqtype=="virus" & viruses$keep_score_all>=1] <- "true positive"
viruses$confusion_matrix[viruses$seqtype!="virus" & viruses$keep_score_all>=1] <- "false positive"
```

```{r}
table(viruses$confusion_matrix, viruses$keep_score_all)
```

```{r}
alluvial_data <- viruses[viruses$confusion_matrix!="true negative",] %>% select(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj, keep_score_all, confusion_matrix) %>%
  count(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj, keep_score_all, confusion_matrix)
```

```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")

ggplot(data = alluvial_data,
       aes(axis1 = as.character(keep_score_vb_cv), 
           axis2 = as.character(keep_score_dvf_cv),
           axis3 = as.character(keep_score_vs2_cv),
           axis4 = as.character(keep_score_cv_kj),
           y = n)) +
  geom_alluvium(aes(fill = confusion_matrix)) +
  geom_stratum() +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum))) +
  scale_fill_manual(name="",
                     values = alpha(pal(4)[c(4,3,1)]), 0.7,
                    labels=c("false negative", "false positive", "true positive")) + 
  scale_x_discrete(limits = c("VIBRANT", "DeepVirFinder", "VirSorter2", "Kaiju"),
                                  expand = c(0.05, 0.05, 0.05, 0.05)) +
  ylab("Number of Sequences") +
  theme_classic()
```

visualizing confusion matrix by taxa
```{r}
confusion_by_taxa <- melt(table(viruses$confusion_matrix, viruses$seqtype))
colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","count")
```

```{r}
ggplot(confusion_by_taxa, aes(x=count, y=seqtype,
                   fill=confusion_matrix,
                   color=confusion_matrix)) +
  geom_bar(stat="identity") +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(4)), 0.5),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(4)), 1),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  xlab("Number of Sequences") +
  ylab("") 
```

# Using NMDS to compare the tools
get matrix want to use for beta diversity
```{r}
virus_score_matrix <- viruses %>% select(keep_score_vb, keep_score_dvf,keep_score_vs2, keep_score_cv, keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj, keep_score_all)

```

```{r}
virus_score_matrix <- virus_score_matrix[rowSums(virus_score_matrix)>0,]
```

```{r}
library(vegan)
```

```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 20", type="regular")

example_NMDS <- metaMDS(t(virus_score_matrix), distance = "bray")

ss <- as.data.frame(example_NMDS$points)
ss$tool <- colnames(virus_score_matrix)

p <- ggplot() + 
  geom_point(data=ss, aes(x=MDS1,y=MDS2, color=tool, fill=tool, shape=tool, size=4)) +
   theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text=element_text(size=14, colour="black"),
        axis.title=element_text(size=16),
        legend.key=element_blank(),
        legend.position="right",
        legend.background = element_blank()) +
  guides(size=FALSE) + xlab("NMDS 1") + ylab("NMDS 2") + 
  coord_equal() +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(nrow(ss))), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(nrow(ss))), 1)) +
  scale_shape_manual(name="",
                     values = c(20,21,21,22,22,23,23,24,24))

p
```

## Looking at False Positives
```{r}
viruses_false_positive <- viruses[viruses$confusion_matrix=="false positive",]
```

```{r}
table(viruses_false_positive$provirus)
```

```{r}
table(viruses_false_positive$vs2type)
```

```{r}
table(viruses_false_positive$percent_viral>50)
```

```{r}
table(viruses_false_positive$hallmark>5)
```

```{r}
table(viruses_false_positive$len.x<=10000)
```

```{r}
table(viruses_false_positive$seqtype)
```