### 2022-10-05 TestingPipeline/Scripts/comparing_rule_sets.Rmd

---
title: "Comparing rule sets"
author: Bridget Hegarty, James Riddell
date: 09-08-2022
output: html_notebook
---
This Rmarkdown file will compare the viral predictions of multiple pipelines 
based on the output of CheckV, DeepVirFinder, Kaiju,VIBRANT, VirSorter, and 
VirSorter2 on multiple training sets of microbial DNA, 
primarily from NCBI. Created from fungal, viral, bacterial, archeael, protist,
and plasmid DNA sequences.

Please reach out to James Riddell (riddell.26@buckeyemail.osu.edu) or
Bridget Hegarty (beh53@case.edu) regarding any issues, or open an issue on github.

```{r setup-library}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)
options(warn=-1)
```

# Inputs
1) For this script's inputs, the user must merge each viral identification 
tool output file into one tab-separated file. 
Replace the filenames with their paths if needed into the variables in the 
'inputs' chunk.

    General filenames for each tool:
        Checkv: quality_summary.tsv
        VIBRANT: VIBRANT_genome_quality_${assembly}.tsv
        Virsorter: VIRSorter_global-phage-signal.csv
        Virsorter2: final-viral-score.tsv
        DeepVirFinder: ${assembly}.fasta_gt2500bp_dvfpred.txt
        Kaiju: ${assembly}.kaiju.names.out

2) Make sure all fasta headers (assembly, contig) are consistent within and
across tools. Chunks for each tool do contain some lines for cleaning up these 
features, but due to their variability it will be the user's responsibility to 
make sure they match across tools.

3) Check each chunk and ensure all columns are accounted for.

4) This script is designed for contigs > 3000 basepairs. It can be modified
to be higher or lower, but going lower will greatly increase the size of the
dataframe and memory usage.

```{r inputs}
# dataset name for file organization and outputs for metrics file
dataset_name <- "testing_set"
# checkV
checkV_path <- "../ToolOutput/testing_sets_checkv_output.tsv"
# VIBRANT
vibrant_path <- "../ToolOutput/testing_sets_vibrant_output.tsv"
# DeepVirFinder
dvf_path <- "../ToolOutput/testing_sets_deepvirfinder_output.tsv"
# Virsorter
vs_path <- "../ToolOutput/testing_sets_virsorter_output.tsv"
# Virsorter2
vs2_path <- "../ToolOutput/testing_sets_virsorter2_output.tsv"
# Kaiju
kj_path <- "../ToolOutput/merged.nreuk.kaiju.names.out"
# KB cutoff
KB_CUTOFF <- 3000
```

# All Viral Contigs

## checkV
```{r}
checkV <- fread(checkV_path, 
                sep="\t",
                header = T, 
                select = c(
                    'Index',
                    'contig_id',
                    'provirus',
                    'completeness',
                    'contamination',
                    'viral_genes',
                    'host_genes',
                    'gene_count',
                    'contig_length',
                    'checkv_quality'
                    )
                ) %>% 
    rename(
        contig = contig_id,
        checkv_provirus = provirus,
        checkv_completeness = completeness,
        checkv_contamination = contamination,
        checkv_viral_genes = viral_genes,
        checkv_host_genes = host_genes,
        checkv_total_genes = gene_count,
        checkv_length = contig_length
        )
checkV$method = 'checkv'
checkV <- separate(checkV, col = contig, into = c("seqtype", "contig"), sep="--")
checkV$contig <- sub("\\.", "_", checkV$contig)
checkV$contig <- sub("\\|", "_", checkV$contig)
```

## VIBRANT
```{r}
vb_c <- fread(vibrant_path,
              header = T,
              sep = "\t",
              select = c(
                  'Index',
                  'scaffold',
                  'type',
                  'Quality'
                  )
              ) %>%
    rename(
        contig = scaffold,
        vibrant_quality = Quality
        )
vb_c$method <- "vibrant"
vb_c$vibrant_prophage <- "No"
vb_c$vibrant_prophage[grep("_fragment_", vb_c$contig)] <- "Yes"
vb_c <- separate(vb_c, col = contig, into = c("seqtype", "contig"), sep="--")
vb_c$contig <- gsub("_fragment_.*", "", vb_c$contig)
vb_c$contig <- sub("\\.", "_", vb_c$contig)
vb_c <- separate(vb_c, col=contig, into="contig", remove=T, sep = " ")
vb_c <- vb_c[!duplicated(vb_c$contig),]
```

## DeepVirFinder
```{r}
dvf_c <- fread(dvf_path,
               header = T,
               sep = "\t",
               select = c(
                   'Index',
                   'name',
                   'score',
                   'pvalue'
                   )
               ) %>% 
    rename(
        contig = name
        )
dvf_c$contig <- sub("\\.", "_", dvf_c$contig)
dvf_c <- separate(dvf_c, col = contig, into = c("seqtype", "contig"), sep="--")
dvf_c$bh_pvalue <- p.adjust(dvf_c$pvalue, method="BH")
```

## VirSorter
```{r}
vs_c <- fread(vs_path,
              select = c(
                  'index',
                  'contig',
                  'category'
                  )
              ) %>% 
    rename(
        Index = index
        )
vs_c$contig <- sub("VIRSorter_", "", vs_c$contig)
vs_c$contig <- sub("-circular", "", vs_c$contig)
vs_c <- separate(vs_c, col = `contig`, into = c("seqtype", "contig"), sep="--")
vs_c$contig <- sub("\\.", "_", vs_c$contig)
vs_c <- vs_c %>% drop_na(contig)
vs_c <- vs_c[!duplicated(vs_c$contig),]
```

## VirSorter2 
```{r}
vs2_c <- fread(vs2_path,
                    header = T,
                    sep = '\t',
                    select = c(
                        'Index',
                        'seqname',
                        'dsDNAphage',
                        'ssDNA',
                        'max_score',
                        'max_score_group',
                        'hallmark',
                        'viral',
                        'cellular'
                    )
               ) %>% 
              separate(
                  col = seqname,
                  into = c("contig", "vs2type"), 
                  sep = "\\|\\|",
                  remove = T
                  )
vs2_c$contig <- sub("\\.", "_", vs2_c$contig)
vs2_c <- separate(vs2_c, col = contig, into = c("seqtype", "contig"), sep="--")
vs2_c <- vs2_c[!duplicated(vs2_c$contig),]
```

## Kaiju 
```{r}
kj_c <- read_tsv(kj_path, col_names = T)
kj_c$Contig <- sub("\\.", "_", kj_c$Contig)
kj_c <- separate(kj_c, col = Contig, into = c("seqtype", "contig"), sep="--")
kj_c <- separate(kj_c, col = Name, into = c("Kaiju_Viral","Kingdom"), sep=";")
kj_c$contig <- sub("\\.", "_", kj_c$contig)
```
Quick glimpse of cellular vs viral sequence classification by Kaiju by assembly
```{r}
table(kj_c$seqtype, kj_c$Kaiju_Viral)
```

## Merging
```{r}
viruses <- full_join(x=checkV, y=kj_c, by = c("Index", "contig", "seqtype"))
viruses <- full_join(x=viruses, y=dvf_c, by = c("Index", "contig", "seqtype"))
viruses <- full_join(x=viruses, y=vb_c, by = c("Index", "contig", "seqtype"))
viruses <- full_join(x=viruses, y=vs_c, by = c("Index", "contig", "seqtype"))
viruses <- full_join(x=viruses, y=vs2_c, by = c("Index", "contig", "seqtype"))
```

Remove contigs not greater than the basepair length cutoff (3000)
```{r}
viruses <- viruses %>% filter(checkv_length > KB_CUTOFF)
```

Use the next code chunk to check if any contigs are missing from checkv.
If yes, then there is a mix up in contig names since checkv should contain
all contig names even if they are identified as non-viral.
```{r}
v_missing <- viruses[is.na(viruses$checkv_uniq_contig),]
```

calculate percent viral/host/unknown
```{r}
viruses$percent_host <- viruses$checkv_host_genes/viruses$checkv_total_genes*100
viruses$percent_viral <- viruses$checkv_viral_genes/viruses$checkv_total_genes*100
viruses$percent_unknown <- 100-(viruses$checkv_host_genes+viruses$checkv_viral_genes)/viruses$checkv_total_genes*100
```

get rid of NAs for downstream processing
```{r}
# checkV
viruses$percent_viral[is.na(viruses$percent_viral)] <- 0
viruses$percent_unknown[is.na(viruses$percent_unknown)] <- 0
# VIBRANT
viruses$vibrant_quality[is.na(viruses$vibrant_quality)] <- 0
# DeepVirFinder
viruses$score[is.na(viruses$score)] <- 0
viruses$bh_pvalue[is.na(viruses$bh_pvalue)] <- 0
# VirSorter2
viruses$viral[is.na(viruses$viral)] <- 0
viruses$hallmark[is.na(viruses$hallmark)] <- 0
#Virsorter
viruses$category[is.na(viruses$category)] <- 0
# Kaiju
viruses$Kaiju_Viral[is.na(viruses$Kaiju_Viral)] <- "unknown"
viruses$Kingdom[is.na(viruses$Kingdom)] <- "unknown"
```

This section defines a viralness score "keep_score" based on the tool classifications. 
A final keep_score above 1 indicates we will keep that sequence and call it viral.

VIBRANT
    Quality == "High Quality Draft": +1
    Quality == "Medium Quality Draft": +1
    Quality == "Low Quality Draft" & provirus == TRUE: +0.5

Virsorter2
    Viral >= 50: +0.5
    Viral >= 0.95: +1
    Hallmark > 2: +1

Virsorter
    category ==  1,2,4,5: +1
    category == 3,6: +0.5

DeepVirFinder:
    Score >= 0.7: +0.5

Kaiju:
    Kaiju_viral = "cellular organisms": -1
    Kaiju_viral = "Viruses": +1

CheckV
    viral_genes == 0 and host_genes >= 1: keep_score = 0
    If 3*viral_genes <= host_genes: keep_score = 0
    If length > 50,000 and hallmark == 0: keep_score = 0
    If %unknown >= 75: +0.5

This script produces visualizations of these combined viral scorings and
includes ecological metrics like alpha diversity.

You can decide which combination is appropriate for them and only need use the
tools appropriate for your data.

```{r getting_viral_set_1}
getting_viral_set_1 <- function(input_seqs,
                                include_vibrant=FALSE, 
                                include_virsorter2=FALSE,
                                include_deepvirfinder=FALSE,
                                include_checkV=FALSE,
                                include_kaiju=FALSE,
                                include_virsorter=FALSE) {

  keep_score <- rep(0, nrow(input_seqs))

  if (include_vibrant) {
    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] + 0.5
  }

  if (include_virsorter2) {
    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.5
    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
  }

  if (include_virsorter) {
    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 1
    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 1
    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5
  }

  if (include_deepvirfinder) {
    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
#   keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 0.5
  }
  if (include_kaiju) {
    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 1
    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 1
  }

  if (include_checkV) {
    keep_score[input_seqs$percent_unknown>=75] <- keep_score[input_seqs$percent_unknown>=75] + 0.5
    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0
    keep_score[input_seqs$contig_length>50000 & input_seqs$hallmark==0] <- 0
  }

  return(keep_score)

}
```

this rule set is based on an original set of rules considered for analyzing the
drinking water virome. upon analyzing the sequences, many false positive viruses
were identified (particularly long sequences with no viral genes)
note: checkV is used as a viral sorting tool
note: VirFinder was used not DeepVirFinder
```{r getting_viral_set_2}
getting_viral_set_2 <- function(input_seqs,
                                include_vibrant=FALSE, 
                                include_virsorter2=FALSE,
                                include_deepvirfinder=FALSE,
                                include_checkV=FALSE,
                                include_virsorter=FALSE) {

  keep_score <- rep(0, nrow(input_seqs))

  if (include_vibrant) {
    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- 
      keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 0.5
    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.3
  }

  if (include_virsorter2) {
    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.3
    keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] <- 
      keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] + 0.5
    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
    }

  if (include_virsorter) {
    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.3
    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.3
  }

  if (include_deepvirfinder) {
    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
    keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 1
  }

  if (include_checkV) {
    keep_score[input_seqs$checkv_quality=="low"] <- keep_score[input_seqs$checkv_quality=="low"] + 0.3
    keep_score[input_seqs$checkv_quality=="medium"] <- keep_score[input_seqs$checkv_quality=="medium"] + 0.5
    keep_score[input_seqs$checkv_quality=="high"] <- keep_score[input_seqs$checkv_quality=="medium"] + 1

    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>1] <- 0
  }

  return(keep_score)

}
```

this is the set of rules used in the drinking water viromes paper
note: VirFinder was used not DeepVirFinder
```{r getting_viral_set_3}
getting_viral_set_3 <- function(input_seqs,
                                include_vibrant=FALSE, 
                                include_virsorter2=FALSE,
                                include_deepvirfinder=FALSE,
                                include_checkV=FALSE,
                                include_virsorter=FALSE) {

  keep_score <- rep(0, nrow(input_seqs))

  if (include_vibrant) {
    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- 
      keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 0.5
    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.3
  }

  if (include_virsorter2) {
    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.3
    keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] <- 
      keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] + 0.5
    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
    }

  if (include_virsorter) {
    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.3
    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.3
  }

  if (include_deepvirfinder) {
    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
    keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 1
  }

  if (include_checkV) {
    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0
    keep_score[input_seqs$contig_length>50000 & input_seqs$hallmark==0] <- 0
  }

  return(keep_score)

}
```

# Assessing performance against the "truth"
note that this is only as accurate as the annotations of the input sequences

this function calculates the precision, recall, and F1 score for each pipeline
```{r}
assess_performance <- function(seqtype, keep_score) {

  truepositive <- rep("not viral", length(seqtype))
  truepositive[seqtype=="virus"] <- "viral"

  #make confusion matrix
  confusion_matrix <- rep("true negative", length(keep_score))
  confusion_matrix[truepositive=="viral" & keep_score<=1] <- "false negative"
  confusion_matrix[truepositive=="viral" & keep_score>=1] <- "true positive"
  confusion_matrix[truepositive=="not viral" & keep_score>=1] <- "false positive"

  precision <- table(confusion_matrix)[4]/(table(confusion_matrix)[4]+table(confusion_matrix)[2])
  recall <- table(confusion_matrix)[4]/(table(confusion_matrix)[4]+table(confusion_matrix)[1])
  F1 <- 2*precision*recall/(precision+recall)

  performance <- c(precision, recall, F1)
  names(performance) <- c("precision", "recall", "F1")

  return(performance)
}
```

combination of tools list
```{r}
combos_list <- read_csv("combinations_list.csv", col_names = T)
```

this function builds a list of all of the combinations that the user wants to 
test. 
In this case, we're comparing the performance of all unique combinations of the 
six tools.
```{r}
build_score_list <- function(input_seqs, combos, pipeline) {
  output <- data.frame(precision=rep(0, nrow(combos)),
                       recall=rep(0, nrow(combos)),
                       F1=rep(0, nrow(combos)))
  for (i in 1:nrow(combos)) {
    if (pipeline==1) {
      keep_score <- getting_viral_set_1(input_seqs, include_vibrant = combos$VIBRANT[i],
                                              include_virsorter = combos$VS[i],
                                              include_virsorter2 = combos$VS2[i],
                                              include_checkV = combos$CheckV[i],
                                              include_kaiju = combos$Kaiju[i],
                                              include_deepvirfinder = combos$DVF[i])
    }
    if (pipeline==2) {
      keep_score <- getting_viral_set_2(input_seqs, include_vibrant = combos$VIBRANT[i],
                                              include_virsorter = combos$VS[i],
                                              include_virsorter2 = combos$VS2[i],
                                              include_checkV = combos$CheckV[i],
                                              include_deepvirfinder = combos$DVF[i])
    }
    if (pipeline==3) {
      keep_score <- getting_viral_set_3(input_seqs, include_vibrant = combos$VIBRANT[i],
                                              include_virsorter = combos$VS[i],
                                              include_virsorter2 = combos$VS2[i],
                                              include_checkV = combos$CheckV[i],
                                              include_deepvirfinder = combos$DVF[i])
    }

    output[i,1:3] <- assess_performance(input_seqs$seqtype, keep_score)

    output$toolcombo[i] <- paste(combos$CheckV[i],combos$DVF[i],
                                 combos$Kaiju[i], combos$VIBRANT[i],
                                 combos$VS[i], combos$VS2[i])
  }

  output[is.na(output)] <- 0

  #return(keep_score)
  return (output)
}
```

## Calculate the performance of each pipeline
first rule set
```{r}
accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*10),
                      precision=rep(0, nrow(combos_list)*10),
                       recall=rep(0, nrow(combos_list)*10),
                       F1=rep(0, nrow(combos_list)*10))
accuracy_scores <- cbind(testing_set_index=rep(1, nrow(combos_list)),
                              build_score_list(viruses[viruses$Index==1], combos_list, pipeline=1))
for (i in 2:10) {
  accuracy_scores <- rbind(accuracy_scores,
                           cbind(testing_set_index=rep(i, nrow(combos_list)),
                              build_score_list(viruses[viruses$Index==i], combos_list, pipeline=1)))
}
```

```{r}
library("stringr")
accuracy_scores$numtools <- str_count(accuracy_scores$toolcombo, "1")
accuracy_scores <- accuracy_scores[order(accuracy_scores$numtools, decreasing=F),]
accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
accuracy_scores$numtools <- as.factor(accuracy_scores$numtools)
```

## Visualize how the precision, recall, and F1 scores change across pipelines.
```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
p2 <- ggplot(accuracy_scores, aes(x=toolcombo, y=F1, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
  ylab("F1 Score")
p2
ggplot(accuracy_scores, aes(x=toolcombo, y=precision)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
  ylab("Precision")
ggplot(accuracy_scores, aes(x=toolcombo, y=recall)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
  ylab("Recall")
```
## Calculate the performance of each pipeline
original rule set from DW paper (didn't use)

```{r}
combos_list_3 <- read_csv("combinations_list_3.csv")
```

```{r}
options(warn=-1)
accuracy_scores_2 <- cbind(testing_set_index=rep(0, nrow(combos_list_3)),
                              build_score_list(viruses[viruses$Index==1], combos_list_3, pipeline=2))
for (i in 2:10) {
  accuracy_scores_2 <- rbind(accuracy_scores_2,
                           cbind(testing_set_index=rep(i, nrow(combos_list_3)),
                              build_score_list(viruses[viruses$Index==i], combos_list_3, pipeline=2)))
}
```

## Visualize how the precision, recall, and F1 scores change across pipelines.
```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
ggplot(accuracy_scores_2, aes(x=toolcombo, y=F1)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (CV, DVF, VB, VS, VS2)") +
  ylab("F1 Score")
ggplot(accuracy_scores_2, aes(x=toolcombo, y=precision)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (CV, DVF, VB, VS, VS2)") +
  ylab("Precision")
ggplot(accuracy_scores_2, aes(x=toolcombo, y=recall)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (CV, DVF, VB, VS, VS2)") +
  ylab("Recall")
```

## Calculate the performance of each pipeline
rule set from DW paper

```{r}
combos_list_3 <- read_csv("combinations_list_3.csv")
```

```{r}
options(warn=-1)
accuracy_scores_3 <- data.frame(testing_set_index=rep(0, nrow(combos_list_3)*10),
                      precision=rep(0, nrow(combos_list_3)*10),
                       recall=rep(0, nrow(combos_list_3)*10),
                       F1=rep(0, nrow(combos_list_3)*10))
accuracy_scores_3 <- cbind(testing_set_index=rep(1, nrow(combos_list_3)),
                              build_score_list(viruses[viruses$Index==1], combos_list_3, pipeline=3))
for (i in 2:10) {
  accuracy_scores_3 <- rbind(accuracy_scores_3,
                           cbind(testing_set_index=rep(i, nrow(combos_list_3)),
                              build_score_list(viruses[viruses$Index==i], combos_list_3, pipeline=3)))
}
```

## Visualize how the precision, recall, and F1 scores change across pipelines.
```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
p2 <- ggplot(accuracy_scores_3, aes(x=toolcombo, y=F1)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (CV, DVF, VB, VS, VS2)") +
  ylab("F1 Score")
p2
ggplot(accuracy_scores_3, aes(x=toolcombo, y=precision)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (CV, DVF, VB, VS, VS2)") +
  ylab("Precision")
ggplot(accuracy_scores_3, aes(x=toolcombo, y=recall)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (CV, DVF, VB, VS, VS2)") +
  ylab("Recall")
```