### 2022-10-28 TestingPipeline/Scripts/testing_set_features_visualization.Rmd
Updating rules with different viral types from VS2.

---
title: "Testing Set Features Visualization"
output: html_notebook
---

This notebook examines the relationship between different features of the data for distinguishing viral from nonviral sequences.

Please reach out to James Riddell (riddell.26@buckeyemail.osu.edu) or
Bridget Hegarty (beh53@case.edu) regarding any issues, or open an issue on github.

```{r setup-library}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)
```

```{r}
viruses <- read_tsv("../IntermediaryFiles/viral_tools_combined.tsv")
```

```{r}
colnames(viruses)
```

```{r}
plot(viruses$RNA, viruses$viral)
plot(viruses$lavidaviridae, viruses$viral)
plot(viruses$NCLDV, viruses$viral)
plot(viruses$ssDNA, viruses$viral)
```

# Important features by sequence type

```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
```

```{r}
ggplot(viruses, aes(x=hallmark, y=checkv_viral_genes)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Number of Hallmark Genes") +
  ylab("Number of Viral Genes")
```

```{r}
ggplot(viruses, aes(x=checkv_host_genes, y=checkv_viral_genes)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Number of Host Genes") +
  ylab("Number of Viral Genes")
```

```{r}
ggplot(viruses, aes(x=percent_unknown, y=percent_viral)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Percentage of Genes Unknown") +
  ylab("Percentage of Genes Viral")
```

```{r}
ggplot(viruses, aes(x=percent_unknown, y=checkv_length)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Percentage of Genes Unknown") +
  ylab("Length of Sequence")
```

```{r}
ggplot(viruses, aes(x=percent_viral, y=checkv_length)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Percentage of Genes Viral") +
  ylab("Length of Sequence") 
```

```{r}
ggplot(viruses, aes(x=hallmark, y=checkv_length)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Number of Hallmark Genes") +
  ylab("Length of Sequence") 
```

```{r}
ggplot(viruses, aes(x=checkv_length, y=checkv_completeness)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Length") +
  ylab("Completeness") 
```

```{r}
ggplot(viruses, aes(x=hallmark, y=checkv_completeness)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Hallmark Genes") +
  ylab("Completeness") 
```

```{r}
table(viruses$seqtype[viruses$checkv_length>50000 & viruses$hallmark==0])/table(viruses$seqtype)
table(viruses$seqtype[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & viruses$checkv_provirus=="No"])/table(viruses$seqtype)
table(viruses$seqtype[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1])/table(viruses$seqtype)

table(viruses$seqtype[viruses$percent_viral>=50])/table(viruses$seqtype)
table(viruses$seqtype[viruses$percent_unknown>=75])/table(viruses$seqtype)
table(viruses$seqtype[viruses$percent_unknown>=75 & viruses$checkv_length<50000])/table(viruses$seqtype)
table(viruses$seqtype[viruses$hallmark>2])/table(viruses$seqtype[viruses$seqtype %in% unique(viruses$seqtype[viruses$hallmark>2])])
```

```{r}
table(viruses$seqtype, viruses$Kaiju_Viral)
```

```{r}
seqdata <- data.frame(seqtype=viruses$seqtype[!duplicated(viruses$contig)])

rownames(seqdata) <- viruses$contig[!duplicated(viruses$contig)]
```

```{r}
library(phyloseq)
```

```{r}
features_table <- viruses[viruses$Index==1,]
features_table <- features_table[,colnames(features_table) %in% c( 
                                              "checkv_viral_genes",
                                              "checkv_host_genes",
                                              "checkv_unknown_genes",
                                              "checkv_length",
                                              "checkv_completeness",
                                              "checkv_total_genes",
                                              "percent_host",
                                              "percent_viral",
                                              "hallmark",
                                              "percent_unknown"
                                              )]

features_table[is.na(features_table)] <- 0
ft_colnames <- colnames(features_table)
features_table <- t(features_table)
rownames(features_table) <- ft_colnames
colnames(features_table) <- rownames(seqdata)

physeq_pooled <- phyloseq(otu_table(features_table, taxa_are_rows = T))
```

```{r}
ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
                          shape="numtools", color="num_viruses") + 
  geom_point(size = 3) +
  theme_bw() +
  geom_label(label=seqdata$toolcombo)

phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
                          shape="numtools", color="num_viruses") + 
  geom_point(size = 3) +
  theme_bw()
```

use of each tuning rule
```{r}
keep_score <- rep(0, nrow(viruses))

keep_score[viruses$hallmark>2] <- keep_score[viruses$hallmark>2] + 1
table(keep_score, viruses$seqtype)
keep_score[viruses$checkv_host_genes>50 & viruses$checkv_provirus=="No"] <- keep_score[viruses$checkv_host_genes>50 & viruses$checkv_provirus=="No"] - 1
table(keep_score, viruses$seqtype)
keep_score[viruses$percent_unknown>=75 & viruses$checkv_length<50000] <- keep_score[viruses$percent_unknown>=75 & viruses$checkv_length<50000] + 0.5
table(keep_score, viruses$seqtype)
keep_score[viruses$percent_viral>=50] <- keep_score[viruses$percent_viral>=50] + 0.5
table(keep_score, viruses$seqtype)
    #keep_score[viruses$hallmark>=(viruses$checkv_viral_genes/5)] <- keep_score[viruses$hallmark>=(viruses$checkv_viral_genes/5)] + 1 #add some ratio
keep_score[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1] <- keep_score[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1] - 1
table(keep_score, viruses$seqtype)
keep_score[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & viruses$checkv_provirus=="No"] <- keep_score[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & viruses$checkv_provirus=="No"] - 1 # consider accounting for provirus designation
table(keep_score, viruses$seqtype)
#    keep_score[(viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes] <- 0 # consider accounting for provirus designation
keep_score[viruses$checkv_length>500000 & viruses$hallmark==0] <- keep_score[viruses$checkv_length>500000 & viruses$hallmark==0] - 1
table(keep_score, viruses$seqtype)
```