### 2022-12-04 TestingPipeline/Scripts/comparing_to_VS2_SOP.Rmd
Testing rule set based on VS2 SOP.

---
title: "Comparing to VS2 SOP"
output: html_notebook
---

# VS2 SOP testing

```{r setup-library}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)
library(pROC)
library("stringr")
```

Import the file that combines the results from each of the tools from running "combining_tool_output.Rmd":
```{r}
viruses <- read_tsv("../IntermediaryFiles/viral_tools_combined.tsv")
```

```{r virsorter2_SOP}
virsorter2_SOP <- function(input_seqs) {
  keep_score <- rep(0, nrow(input_seqs))
  # Keep1: viral_gene >0
  keep_score[input_seqs$checkv_viral_genes > 0] <- 1

  # Keep2: viral_gene =0 AND (host_gene =0 OR score >=0.95 OR hallmark >2)
  keep_score[(input_seqs$checkv_viral_genes == 0) & ((input_seqs$checkv_host_genes == 0) | (input_seqs$max_score >=0.95) | (input_seqs$hallmark > 2))] <- 1

  # Manual check: (NOT in Keep1 OR Keep2) AND viral_gene =0 AND host_gene =1 AND length >=10kb
  keep_score[!(input_seqs$checkv_viral_genes > 0) & !((input_seqs$checkv_viral_genes == 0) & ((input_seqs$checkv_host_genes == 0) | (input_seqs$max_score >=0.95) | (input_seqs$hallmark > 2))) & (input_seqs$checkv_viral_genes == 0) & (input_seqs$checkv_host_genes == 1) & (input_seqs$checkv_length > 10000)] <- 0.5

  # Discard: the rest
  # implied by starting with zero matrix, unchanged will stay zero

  return(keep_score)
}

```

```{r}
viruses$keep_score_virsorter2_SOP <- virsorter2_SOP(viruses)
```

```{r}
viruses$confusion_matrix_vs2_sop <- "true negative"
viruses$confusion_matrix_vs2_sop[viruses$seqtype=="virus" & viruses$keep_score_virsorter2_SOP<1] <- "false negative"
viruses$confusion_matrix_vs2_sop[viruses$seqtype=="virus" & viruses$keep_score_virsorter2_SOP>=1] <- "true positive"
viruses$confusion_matrix_vs2_sop[viruses$seqtype!="virus" & viruses$keep_score_virsorter2_SOP>=1] <- "false positive"

viruses$accurate_vs2_SOP <- "false"
viruses$accurate_vs2_SOP[grep("true", viruses$confusion_matrix_vs2_sop)] <- "true"
```

accuracy:
```{r}
length(grep("true", viruses$accurate_vs2_SOP))/nrow(viruses)
```

precision
```{r}
length(grep("true positive", viruses$confusion_matrix_vs2_sop))/length(grep("positive", viruses$confusion_matrix_vs2_sop))
```

recall
```{r}
length(grep("true positive", viruses$confusion_matrix_vs2_sop))/(length(grep("true positive", viruses$confusion_matrix_vs2_prsop))+length(grep("false negative", viruses$confusion_matrix_vs2_sop)))
```