### 2023-02-14 TestingPipeline/Scripts/combining_tool_output-fragments.Rmd

---
title: "Combining Tool Output - Fragments"
output: html_notebook
---

This Rmarkdown file combines the output of CheckV, DeepVirFinder, Kaiju,
VIBRANT, VirSorter, and VirSorter2 on multiple training sets of microbial DNA, 
primarily from NCBI. Created from fungal, viral, bacterial, archeael, protist,
and plasmid DNA sequences

Please reach out to James Riddell (riddell.26@buckeyemail.osu.edu) or
Bridget Hegarty (beh53@case.edu) regarding any issues, or open an issue on github.

```{r setup-library}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)
```

# Inputs
1) For this script's inputs, the user must merge each viral identification 
tool output file into one tab-separated file. 
Replace the filenames with their paths if needed into the variables in the 
'inputs' chunk.

    General filenames for each tool:
        Checkv: quality_summary.tsv
        VIBRANT: VIBRANT_genome_quality_${assembly}.tsv
        Virsorter: VIRSorter_global-phage-signal.csv
        Virsorter2: final-viral-score.tsv
        DeepVirFinder: ${assembly}.fasta_gt2500bp_dvfpred.txt
        Kaiju: ${assembly}.kaiju.names.out

2) Make sure all fasta headers (assembly, contig) are consistent within and
across tools. Chunks for each tool do contain some lines for cleaning up these 
features, but due to their variability it will be the user's responsibility to 
make sure they match across tools.

3) Check each chunk and ensure all columns are accounted for.

4) This script is designed for contigs > 3000 basepairs. It can be modified
to be higher or lower, but going lower will greatly increase the size of the
dataframe and memory usage.

5) This script only keeps each fragment once. This has two purposes: (1) it 
deals with the duplicate entries for fragments labeled as "complete" (the second
entry gives the quality); (2) it keeps only one instance of sequences that have
multiple viral fragments on them.

```{r inputs}
# dataset name for file organization and outputs for metrics file
dataset_name <- "testing_set-fragments"
# checkV
checkV_path <- "../ToolOutput/Fragments_Results/merged_checkv_assemblies.tsv"
# VIBRANT
vibrant_path <- "../ToolOutput/Fragments_Results/merged_vibrant_assemblies.tsv"
# DeepVirFinder
dvf_path <- "../ToolOutput/Fragments_Results/merged_dvf_assemblies.tsv"
# Virsorter
vs_path <- "../ToolOutput/Fragments_Results/merged_virsorter_assemblies.tsv"
# Virsorter2
vs2_path <- "../ToolOutput/Fragments_Results/merged_vs2_assemblies.tsv"
# Kaiju
kj_path <- "../ToolOutput/Fragments_Results/merged_kaiju_assemblies.tsv"
# KB cutoff
KB_CUTOFF <- 3000
```

# All Viral Contigs

## checkV
```{r}
checkV <- fread(checkV_path, 
                sep="\t",
                header = T, 
                select = c(
                    'index',
                    'contig_id',
                    'provirus',
                    'completeness',
                    'contamination',
                    'viral_genes',
                    'host_genes',
                    'gene_count',
                    'contig_length',
                    'checkv_quality'
                    )
                ) %>% 
    rename(
        Index = index,
        contig = contig_id,
        checkv_provirus = provirus,
        checkv_completeness = completeness,
        checkv_contamination = contamination,
        checkv_viral_genes = viral_genes,
        checkv_host_genes = host_genes,
        checkv_total_genes = gene_count,
        checkv_length = contig_length
        )
checkV$method = 'checkv'
checkV <- separate(checkV, col = contig, into = c("seqtype", "contig"), sep="--")
checkV$contig <- sub("\\.", "_", checkV$contig)
checkV$contig <- sub("\\|", "_", checkV$contig)
checkV <- checkV[!duplicated(paste(checkV$contig, checkV$Index)),]
```

## VIBRANT
```{r}
vb_c <- fread(vibrant_path,
              header = T,
              sep = "\t",
              select = c(
                  'index',
                  'scaffold',
                  'type',
                  'Quality'
                  )
              ) %>%
    rename(
        Index = index,
        contig = scaffold,
        vibrant_quality = Quality
        )
vb_c$method <- "vibrant"
vb_c$vibrant_prophage <- "No"
vb_c$vibrant_prophage[grep("_fragment_", vb_c$contig)] <- "Yes"
vb_c <- separate(vb_c, col = contig, into = c("seqtype", "contig"), sep="--")
vb_c$contig <- gsub("_fragment_.*", "", vb_c$contig)
vb_c$contig <- sub("\\.", "_", vb_c$contig)
vb_c <- separate(vb_c, col=contig, into="contig", remove=T, sep = " ")
vb_c <- vb_c[!duplicated(paste(vb_c$contig, vb_c$Index)),]
```

## DeepVirFinder
```{r}
dvf_c <- fread(dvf_path,
               header = T,
               sep = "\t",
               select = c(
                   'index',
                   'name',
                   'score',
                   'pvalue'
                   )
               ) %>% 
    rename(
        Index = index,
        contig = name
        )
dvf_c$contig <- sub("\\.", "_", dvf_c$contig)
dvf_c <- separate(dvf_c, col = contig, into = c("seqtype", "contig"), sep="--")
dvf_c$bh_pvalue <- p.adjust(dvf_c$pvalue, method="BH")
dvf_c <- dvf_c[!duplicated(paste(dvf_c$contig, dvf_c$Index)),]
```

## VirSorter
```{r}
vs_c <- fread(vs_path,
              select = c(
                  '## Contig_id',
                  'Category'
                  )
              ) %>% 
    rename(
        contig = '## Contig_id',
        category = Category,
        )
vs_c <- vs_c[!grep("#", vs_c$contig),]
vs_c <- separate(vs_c, col = `contig`, into = c("Index", "type", "contig"), sep="--")
vs_c$seqtype <- sub("VIRSorter_", "", vs_c$type)
vs_c$contig <- sub("\\.", "_", vs_c$contig)
vs_c <- vs_c %>% drop_na(contig)
vs_c <- vs_c[!duplicated(paste(vs_c$contig, vs_c$Index)),]
vs_c$Index <- as.numeric(vs_c$Index)
vs_c$category <- as.numeric(vs_c$category)
```

## VirSorter2 
```{r}
vs2_c <- fread(vs2_path,
                    header = T,
                    sep = '\t',
                    select = c(
                        'index',
                        'seqname',
                        'dsDNAphage',
                        'ssDNA',
                        'NCLDV',
                        'RNA',
                        'lavidaviridae',
                        'max_score',
                        'max_score_group',
                        'hallmark',
                        'viral',
                        'cellular'
                    )
               ) %>% 
            rename(
                Index = index
            ) %>%
              separate(
                  col = seqname,
                  into = c("contig", "vs2type"), 
                  sep = "\\|\\|",
                  remove = T
                  )
vs2_c$contig <- sub("\\.", "_", vs2_c$contig)
vs2_c <- separate(vs2_c, col = contig, into = c("seqtype", "contig"), sep="--")
vs2_c <- vs2_c[!duplicated(paste(vs2_c$contig, vs2_c$Index)),]
```

## Kaiju 
```{r}
kj_c <- read_tsv(kj_path, col_names = T)
colnames(kj_c)[2] <- "Contig"
colnames(kj_c)[ncol(kj_c)] <- "Index"
kj_c$Contig <- sub("\\.", "_", kj_c$Contig)
kj_c <- separate(kj_c, col = Contig, into = c("seqtype", "contig"), sep="--")
kj_c <- separate(kj_c, col = taxonomy, into = c("Kaiju_Viral","Kingdom"), sep=";")
kj_c$contig <- sub("\\.", "_", kj_c$contig)
kj_c <- kj_c[!duplicated(paste(kj_c$contig, kj_c$Index)),]
```

## Merging
```{r}
viruses <- full_join(x=checkV, y=kj_c, by = c("Index", "contig", "seqtype"))
viruses <- full_join(x=viruses, y=dvf_c, by = c("Index", "contig", "seqtype"))
viruses <- full_join(x=viruses, y=vb_c, by = c("Index", "contig", "seqtype"))
viruses <- full_join(x=viruses, y=vs_c, by = c("Index", "contig", "seqtype"))
viruses <- full_join(x=viruses, y=vs2_c, by = c("Index", "contig", "seqtype"))
```

Remove contigs not greater than the basepair length cutoff (3000)
```{r}
viruses <- viruses %>% filter(checkv_length > KB_CUTOFF)
```

Use the next code chunk to check if any contigs are missing from checkv.
If yes, then there is a mix up in contig names since checkv should contain
all contig names even if they are identified as non-viral.
```{r}
v_missing <- viruses[is.na(viruses$checkv_uniq_contig),]
```

calculate percent viral/host/unknown
```{r}
viruses$percent_host <- viruses$checkv_host_genes/viruses$checkv_total_genes*100
viruses$percent_viral <- viruses$checkv_viral_genes/viruses$checkv_total_genes*100
viruses$percent_unknown <- 100-(viruses$checkv_host_genes+viruses$checkv_viral_genes)/viruses$checkv_total_genes*100
```

get rid of NAs for downstream processing
```{r}
# checkV
viruses$percent_viral[is.na(viruses$percent_viral)] <- 0
viruses$percent_unknown[is.na(viruses$percent_unknown)] <- 0
viruses$checkv_completeness[is.na(viruses$checkv_completeness)] <- 0

# VIBRANT
viruses$vibrant_quality[is.na(viruses$vibrant_quality)] <- 0

# DeepVirFinder
viruses$score[is.na(viruses$score)] <- 0
viruses$bh_pvalue[is.na(viruses$bh_pvalue)] <- 0
viruses$pvalue[is.na(viruses$pvalue)] <- 0

# VirSorter2
viruses$max_score[is.na(viruses$max_score)] <- 0
viruses$viral[is.na(viruses$viral)] <- 0
viruses$hallmark[is.na(viruses$hallmark)] <- 0
viruses$RNA[is.na(viruses$RNA)] <- 0
viruses$lavidaviridae[is.na(viruses$lavidaviridae)] <- 0
viruses$NCLDV[is.na(viruses$NCLDV)] <- 0
viruses$ssDNA[is.na(viruses$ssDNA)] <- 0
viruses$dsDNAphage[is.na(viruses$dsDNAphage)] <- 0
viruses$max_score[is.na(viruses$max_score)] <- 0

#Virsorter
viruses$category[is.na(viruses$category)] <- 0

# Kaiju
viruses$Kaiju_Viral[is.na(viruses$Kaiju_Viral)] <- "unknown"
viruses$Kingdom[is.na(viruses$Kingdom)] <- "unknown"
viruses$kaiju_match_ratio <- viruses$len/viruses$checkv_length

#adding a combined provirus category
viruses$provirus <- FALSE
viruses$provirus[viruses$checkv_provirus=="Yes" | viruses$vibrant_prophage=="Yes" | viruses$category==4] <- TRUE
```

```{r}
write_tsv(viruses, "../IntermediaryFiles/viral_tools_combined-fragments.tsv")
```