### 2023-03-23 TestingPipeline/Scripts/examining_false_positives.Rmd
Updating figures with current rules.

---
title: "Examining False Negatives"
author: Bridget Hegarty, James Riddell
date: 10-30-2022
output: html_notebook
---
This Rmarkdown file examines the false positives from the output of CheckV, DeepVirFinder, Kaiju,
VIBRANT, VirSorter, and VirSorter2 on multiple training sets of microbial DNA, 
primarily from NCBI. Created from fungal, viral, bacterial, archeael, protist,
and plasmid DNA sequences

Please reach out to James Riddell (riddell.26@buckeyemail.osu.edu) or
Bridget Hegarty (beh53@case.edu) regarding any issues, or open an issue on github.

```{r setup-library}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)
library(pROC)
```

Import the file that combines the results from each of the tools from running "combining_tool_output.Rmd":
```{r}
viruses <- read_tsv("../IntermediaryFiles/viral_tools_combined.tsv")
```

```{r}
viruses$kaiju_match_ratio <- viruses$len/viruses$checkv_length
```

This section defines a viralness score "keep_score" based on the tool classifications. 
A final keep_score above 1 indicates we will keep that sequence and call it viral.

This script produces visualizations of these combined viral scorings and
includes ecological metrics like alpha diversity.

You can decide which combination is appropriate for them and only need use the
tools appropriate for your data.

```{r getting_viral_set_1}
getting_viral_set_1 <- function(input_seqs,
                                include_vibrant=FALSE, 
                                include_virsorter2=FALSE,
                                include_deepvirfinder=FALSE,
                                include_tuning_viral=FALSE,
                                tv_1=T, tv_2=T, tv_3=T, tv_4=T,
                                include_tuning_not_viral=FALSE,
                                ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
                                include_virsorter=FALSE) {

  keep_score <- rep(0, nrow(input_seqs))

  if (include_vibrant) {
    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] + 0.5
  }

  if (include_virsorter2) {
    keep_score[input_seqs$max_score>=0.50] <- keep_score[input_seqs$max_score>=0.50] + 0.5
    keep_score[input_seqs$max_score>=0.95] <- keep_score[input_seqs$max_score>=0.95] + 0.5  
  }

  if (include_virsorter) {
    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5 
    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5 
  }

  if (include_deepvirfinder) {
     # add if DVF calls viral
    keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
    keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
  }

  if (include_tuning_viral) {
    keep_score <- tuning_viral(current_score=keep_score,
                               ins=input_seqs,
                         include_tv_1=tv_1,
                         include_tv_2=tv_2,
                         include_tv_3=tv_3,
                         include_tv_4=tv_4)
  }

  if (include_tuning_not_viral) {
    # tuning removal 
    keep_score <- tuning_not_viral(current_score=keep_score,
                               ins=input_seqs,
                         include_tv_1=ntv_1,
                         include_tv_2=ntv_2,
                         include_tv_3=ntv_3,
                         include_tv_4=ntv_4,
                         include_tv_5=ntv_5)

  }

  return(keep_score)

}

tuning_viral <- function(current_score, ins, 
                         include_tv_1=T,
                         include_tv_2=T,
                         include_tv_3=T,
                         include_tv_4=T) {
    # tuning addition
  if (include_tv_1) {
    current_score[ins$hallmark>2] <- current_score[ins$hallmark>2] + 1
  }
  if (include_tv_2) {
    current_score[ins$Kaiju_Viral=="Viruses" & ins$kaiju_match_ratio>=0.3] <- current_score[ins$Kaiju_Viral=="Viruses" & ins$kaiju_match_ratio>=0.3] + 1
  }
    #note: had tried pulling out this rule, but better recall without sacrificing precision with it in
    #note: having kaiju removal rule didn't help
  if (include_tv_3) {
    current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] <- current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] + 0.5
  }
  if (include_tv_4) {
    current_score[ins$viral>=50 | ins$percent_viral>=50] <- current_score[ins$viral>=50 | ins$percent_viral>=50] + 0.5 
  }
  return(current_score)
}

tuning_not_viral <- function(current_score, ins, 
                         include_tv_1=T,
                         include_tv_2=T,
                         include_tv_3=T,
                         include_tv_4=T,
                         include_tv_5=F) {
    # tuning addition
  if (include_tv_1) {
    current_score[(ins$checkv_host_genes>50) & !ins$provirus] <- 0
  }
  if (include_tv_2) {
    current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] <- current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] - 1
  }
  if (include_tv_3) {
    current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] <- current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] - 1
  }
  if (include_tv_4) {
    current_score[ins$checkv_length>500000 & ins$hallmark<=1] <- 0
  }

  if (include_tv_5) {
    # remove if DVF calls it not viral
    current_score[ins$score<=0.7 & ins$pvalue<=0.05] <- current_score[ins$score<=0.7 & ins$pvalue<=0.05] - 1 
  }

  return(current_score)
}
```

```{r}
viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)
```

```{r}
viruses$confusion_matrix_high_MCC <- "true negative"
viruses$confusion_matrix_high_MCC[viruses$seqtype=="virus" & viruses$keep_score_high_MCC<1] <- "false negative"
viruses$confusion_matrix_high_MCC[viruses$seqtype=="virus" & viruses$keep_score_high_MCC>=1] <- "true positive"
viruses$confusion_matrix_high_MCC[viruses$seqtype!="virus" & viruses$keep_score_high_MCC>=1] <- "false positive"
```

getting a list of false positive bacteria and plasmids from testing set 1
```{r}
fp_bacteria_list <- viruses$contig[viruses$Index==1 & 
                                     viruses$seqtype=="bacteria" &
                                     viruses$confusion_matrix_high_MCC=="false positive"]

fp_plasmid_list <- viruses$contig[viruses$Index==1 & 
                                     viruses$seqtype=="plasmid" &
                                     viruses$confusion_matrix_high_MCC=="false positive"]

tp_virus_list <- viruses$contig[viruses$Index==1 & 
                                     viruses$seqtype=="virus" &
                                     viruses$confusion_matrix_high_MCC=="true positive"]
```

vibrant annotations
```{r}
annotations_vb <- read_tsv("../ToolOutput/VIBRANT_annotations_metagenomic_testing_set_1.tsv")
```
```{r}
annotations_vb <- separate(annotations_vb, col="scaffold", into=c("seqtype", "contig"),
                           sep="--")
annotations_vb$contig <- sub("\\.", "_", annotations_vb$contig)
```

```{r}
head(annotations_vb)
```
```{r}
table(annotations_vb$contig %in% fp_bacteria_list)
table(annotations_vb$contig %in% fp_plasmid_list)
```

```{r}
annotations_vb_fp <- rbind(annotations_vb[annotations_vb$contig %in% fp_bacteria_list,],
                           annotations_vb[annotations_vb$contig %in% fp_plasmid_list,],
                           annotations_vb[annotations_vb$contig %in% tp_virus_list,])
```

```{r}
annotations_vb_fp_vog_count <- annotations_vb_fp %>%
  select(contig, VOG, seqtype) %>%
  group_by(contig, seqtype) %>%
  summarize(num_vog=table(is.na(VOG))[[1]],
            prop_vog=table(is.na(VOG))[[1]]/length(VOG),
            num_orfs=length(VOG))
```

```{r}
annotations_vb_fp_vog_count$confusion_matrix <- "false positive"
annotations_vb_fp_vog_count$confusion_matrix[annotations_vb_fp_vog_count$seqtype=="virus"] <- "true positive"
```

```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
```

```{r}
ggplot(annotations_vb_fp_vog_count, 
       aes(x=seqtype, y=prop_vog, 
           color=seqtype, fill=seqtype)) +
  geom_boxplot() +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  ylab("Proportion Genes VOGs") +
  xlab("Sequence Type") +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(3)), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(3)), 1)) +
  facet_wrap(~confusion_matrix, scales="free_x")
```
```{r}
t.test(annotations_vb_fp_vog_count$prop_vog[annotations_vb_fp_vog_count$seqtype=="bacteria"],
       annotations_vb_fp_vog_count$prop_vog[annotations_vb_fp_vog_count$seqtype=="virus"])
```
```{r}
t.test(annotations_vb_fp_vog_count$prop_vog[annotations_vb_fp_vog_count$seqtype=="plasmid"],
       annotations_vb_fp_vog_count$prop_vog[annotations_vb_fp_vog_count$seqtype=="virus"])
```

```{r}
ggplot(annotations_vb_fp_vog_count, 
       aes(x=num_vog, y=prop_vog, 
           color=seqtype, fill=seqtype)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  ylab("Proportion Genes VOGs") +
  xlab("Number of VOGs") +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(3)), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(3)), 1))
```

# Thinking about short protists
```{r}
viruses[viruses$confusion_matrix_high_MCC=="false positive" & viruses$seqtype=="protist",]

table(viruses$confusion_matrix_high_MCC[viruses$seqtype=="protist"], 
      viruses$max_score_group[viruses$seqtype=="protist"])

table(viruses$Kingdom[viruses$seqtype=="protist" &  viruses$confusion_matrix_high_MCC=="false positive"], 
      viruses$max_score_group[viruses$seqtype=="protist" & viruses$confusion_matrix_high_MCC=="false positive"])

table(viruses$percent_viral[viruses$seqtype=="protist" &  viruses$confusion_matrix_high_MCC=="false positive"]>20, 
      viruses$max_score_group[viruses$seqtype=="protist" & viruses$confusion_matrix_high_MCC=="false positive"])
```

```{r}
ggplot(viruses, aes(y=percent_viral, x=confusion_matrix_high_MCC)) +
  geom_boxplot() +
  geom_point() +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") 
```
```{r}
table(viruses$confusion_matrix_high_MCC[viruses$checkv_length>=5000])
```

# Just looking at the protists getting wrong

```{r}
viruses$kaiju_match_ratio <- viruses$len/viruses$checkv_length
```

```{r}
ggplot(viruses, aes(y=kaiju_match_ratio, x=seqtype)) +
  geom_boxplot() +
  geom_point() +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") 
```