### 2023-03-23 TestingPipeline/Scripts/comparing_to_VS2_SOP.Rmd

---
title: "Comparing to VS2 SOP"
output: html_notebook
---

# VS2 SOP testing of version 2 and 3

```{r setup-library}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)
library(pROC)
library("stringr")
```

Import the file that combines the results from each of the tools from running "combining_tool_output.Rmd":
```{r}
viruses <- read_tsv("../IntermediaryFiles/viral_tools_combined.tsv")
```

```{r virsorter2_SOP v2}
virsorter2_SOP2 <- function(input_seqs) {
  keep_score <- rep(0, nrow(input_seqs))
  # Keep: viral_gene >0 OR score >=0.95 OR hallmark >2 OR (viral_gene =0 AND host_gene =0)
  keep_score[(input_seqs$checkv_viral_genes > 0) |
             (input_seqs$max_score >=0.95) |
             (input_seqs$hallmark > 2) |
             ((input_seqs$checkv_viral_genes == 0) & (input_seqs$checkv_host_genes == 0))] <- 1
  # discard viral_gene =0 AND host_gene >1 OR (viral_gene =0 AND host_gene =1 AND length <10kb)
  keep_score[(input_seqs$checkv_viral_genes == 0) & 
             (input_seqs$checkv_host_genes > 1) |
             ((input_seqs$checkv_viral_genes == 0) & (input_seqs$checkv_host_genes == 1) & (input_seqs$checkv_length < 10000))] <- 0
  # Manual check: (NOT in Keep1 OR Keep2) AND viral_gene =0 AND host_gene =1 AND length >=10kb
  keep_score[(input_seqs$checkv_viral_genes == 0) & (input_seqs$checkv_host_genes == 1) & (input_seqs$checkv_length >= 10000)] <- 0.5
  # implied by starting with zero matrix, unchanged will stay zero

  return(keep_score)
}

```

```{r virsorter2_SOP v3}
virsorter2_SOP3 <- function(input_seqs) {
  keep_score <- rep(0, nrow(input_seqs))
  # Keep1: viral_gene >0
  keep_score[input_seqs$checkv_viral_genes > 0] <- 1

  # Keep2: viral_gene =0 AND (host_gene =0 OR score >=0.95 OR hallmark >2)
  keep_score[(input_seqs$checkv_viral_genes == 0) & ((input_seqs$checkv_host_genes == 0) | (input_seqs$max_score >=0.95) | (input_seqs$hallmark > 2))] <- 0.75

  # Manual check: (NOT in Keep1 OR Keep2) AND viral_gene =0 AND host_gene =1 AND length >=10kb
  keep_score[!(input_seqs$checkv_viral_genes > 0) & !((input_seqs$checkv_viral_genes == 0) & ((input_seqs$checkv_host_genes == 0) | (input_seqs$max_score >=0.95) | (input_seqs$hallmark > 2))) & (input_seqs$checkv_viral_genes == 0) & (input_seqs$checkv_host_genes == 1) & (input_seqs$checkv_length > 10000)] <- 0.5

  # Discard: the rest
  # implied by starting with zero matrix, unchanged will stay zero

  return(keep_score)
}

```

```{r}
viruses$keep_score_virsorter2_SOP3 <- virsorter2_SOP3(viruses)
viruses$keep_score_virsorter2_SOP2 <- virsorter2_SOP2(viruses)
```

```{r}
viruses$confusion_matrix_vs2_sop2 <- "true negative"
viruses$confusion_matrix_vs2_sop2[viruses$seqtype=="virus" & viruses$keep_score_virsorter2_SOP2<1] <- "false negative"
viruses$confusion_matrix_vs2_sop2[viruses$seqtype=="virus" & viruses$keep_score_virsorter2_SOP2>=1] <- "true positive"
viruses$confusion_matrix_vs2_sop2[viruses$seqtype!="virus" & viruses$keep_score_virsorter2_SOP2>=1] <- "false positive"

viruses$accurate_vs2_SOP2 <- "false"
viruses$accurate_vs2_SOP2[grep("true", viruses$confusion_matrix_vs2_sop2)] <- "true"
```

```{r}
viruses$confusion_matrix_vs2_sop3 <- "true negative"
viruses$confusion_matrix_vs2_sop3[viruses$seqtype=="virus" & viruses$keep_score_virsorter2_SOP3<1] <- "false negative"
viruses$confusion_matrix_vs2_sop3[viruses$seqtype=="virus" & viruses$keep_score_virsorter2_SOP3>=1] <- "true positive"
viruses$confusion_matrix_vs2_sop3[viruses$seqtype!="virus" & viruses$keep_score_virsorter2_SOP3>=1] <- "false positive"

viruses$accurate_vs2_SOP3 <- "false"
viruses$accurate_vs2_SOP3[grep("true", viruses$confusion_matrix_vs2_sop3)] <- "true"
```

```{r}
# investigate keep2
keep2 = viruses[viruses$keep_score_virsorter2_SOP3 == 0.75,]
nrow(keep2)
```
```{r}
length(grep("virus", keep2$seqtype))
```

assume keep2 (0.75) and manual curation (0.5) has 100% accuracy to avoid having to check with DRAM v since it represents a small portion of the sequences

```{r}
viruses$confusion_matrix_vs2_keep2 <- "true negative"
viruses$confusion_matrix_vs2_keep2[viruses$seqtype=="virus" & viruses$keep_score_virsorter2_SOP3<1] <- "false negative"
viruses$confusion_matrix_vs2_keep2[viruses$seqtype=="virus" & viruses$keep_score_virsorter2_SOP3>=1] <- "true positive"
viruses$confusion_matrix_vs2_keep2[viruses$seqtype!="virus" & viruses$keep_score_virsorter2_SOP3>=1] <- "false positive"
viruses$confusion_matrix_vs2_keep2[viruses$keep_score_virsorter2_SOP3==0.75] <- "true positive"
viruses$confusion_matrix_vs2_keep2[viruses$keep_score_virsorter2_SOP3==0.5] <- "true positive"

viruses$accurate_vs2_keep2 <- "false"
viruses$accurate_vs2_keep2[grep("true", viruses$confusion_matrix_vs2_keep2)] <- "true"
```

accuracy:
```{r}
length(grep("true", viruses$accurate_vs2_SOP2))/nrow(viruses)
```
```{r}
length(grep("true", viruses$accurate_vs2_SOP3))/nrow(viruses)
```

precision
```{r}
length(grep("true positive", viruses$confusion_matrix_vs2_sop2))/length(grep("positive", viruses$confusion_matrix_vs2_sop2))
```
```{r}
length(grep("true positive", viruses$confusion_matrix_vs2_sop3))/length(grep("positive", viruses$confusion_matrix_vs2_sop3))
```
```{r}
length(grep("true positive", viruses$confusion_matrix_vs2_keep2))/length(grep("positive", viruses$confusion_matrix_vs2_keep2))
```

recall
```{r}
length(grep("true positive", viruses$confusion_matrix_vs2_sop2))/(length(grep("true positive", viruses$confusion_matrix_vs2_sop2))+length(grep("false negative", viruses$confusion_matrix_vs2_sop2)))
```
```{r}
length(grep("true positive", viruses$confusion_matrix_vs2_sop3))/(length(grep("true positive", viruses$confusion_matrix_vs2_sop3))+length(grep("false negative", viruses$confusion_matrix_vs2_sop3)))
```
```{r}
length(grep("true positive", viruses$confusion_matrix_vs2_keep2))/(length(grep("true positive", viruses$confusion_matrix_vs2_keep2))+length(grep("false negative", viruses$confusion_matrix_vs2_keep2)))
```