### 2023-05-27 EnvironmentalData/CrossEnvironmentComparisons/comparing_rule_sets_across_environments.Rmd
updated x axis to match MCC graph
Also added a csv that has num_viruses without drinking water (being loaded from drinking water folder)

---
title: "Comparing rule sets across environments"
output: html_notebook
---

```{r setup-library}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)
```

```{r}
env_path <- '/Users/riddellj/Documents/Research/VSTE/VSTE_Old_Pull/EnvironmentalData/'
viruses_tara <- read_csv(paste(env_path, "tara_num_viruses.csv", sep=""))
viruses_tara$environment <- "global_oceans"
viruses_le <- read_csv(paste(env_path, "le_num_viruses.csv", sep=""))
viruses_le$environment <- "Lake_Erie"
viruses_lm <- read_csv(paste(env_path, "lake_michigan_num_viruses.csv", sep=""))
viruses_lm$environment <- "Lake_Michigan"
viruses_ww <- read_csv(paste(env_path, "wastewater_num_viruses.csv", sep=""))
viruses_ww$environment <- "wastewater"
viruses_dw <- read_csv(paste(env_path, "drinking_water_uknl_num_viruses.csv", sep=""))

viruses_rest <- rbind(viruses_tara, viruses_le, viruses_lm, viruses_ww)
write.csv(viruses_rest, 'env_rest_num_viruses.csv')
```

only keeping chlorinated drinking water samples
```{r}
viruses_dw <- viruses_dw[viruses_dw$assembly %in% c("Birmingham_Maria",
                                                       "Cardiff_Maria", "Dundee_Clat",
                                                       "Dundee_Maria",
                                                       "Exeter_Maria", "Glasgow_Bal",
                                                       "Glasgow_Mil", "London_Maria"),
                           ]
```

```{r}
viruses_dw$environment <- "drinking_water"

# viruses_rest$environment[viruses_rest$environment=="waste_water"] <- "wastewater"
# viruses_rest$environment[viruses_rest$environment=="lake_eerie"] <- "Lake_Erie"
# viruses_rest$environment[viruses_rest$environment=="lake_michigan"] <- "Lake_Michigan"
# viruses_rest$environment[viruses_rest$environment=="tara"] <- "global_oceans"

viruses_all <- rbind(viruses_dw, viruses_rest)
```

recoding AJ's samples to be able to separate virome from metagenome (and remove others)
```{r}
viruses_all$environment2 <- viruses_all$environment
viruses_all$environment2[viruses_all$assembly %in% c("49639", "49640", "49641", "49642")] <- "Lake_Erie_virome"
viruses_all$environment2[viruses_all$assembly %in% c("49613", "49614", "49618", "49621",
                                                     "49624", "49628", "49632", "49638")] <- "Lake_Erie_metagenome"
viruses_all <- viruses_all[viruses_all$environment2!="Lake_Erie",]
```

```{r}
viruses_all$environment2 <- factor(viruses_all$environment2,
                                    levels = c("drinking_water", "Lake_Erie_metagenome", 
                                               "global_oceans",
                                               "wastewater", "Lake_Erie_virome", 
                                               "Lake_Michigan"))
```

```{r}
viruses_all$best_coloring <- "others"
# 
viruses_all$best_coloring[viruses_all$toolcombo=="1 1 1 1 1 1"] <- "all"
# high recall
viruses_all$best_coloring[viruses_all$toolcombo=="1 1 0 1 1 1"] <- "high recall"
viruses_all$best_coloring[viruses_all$toolcombo=="1 1 0 0 1 1"] <- "high recall"
viruses_all$best_coloring[viruses_all$toolcombo=="1 1 0 0 0 1"] <- "high recall"
viruses_all$best_coloring[viruses_all$toolcombo=="1 0 0 0 1 1"] <- "high recall"
viruses_all$best_coloring[viruses_all$toolcombo=="1 1 0 1 0 1"] <- "high recall"
# VS2 Combo
viruses_all$best_coloring[viruses_all$toolcombo=="1 0 1 0 0 1"] <- "VS2 combo"
# High Precision
viruses_all$best_coloring[viruses_all$toolcombo=="0 0 0 0 0 1"] <- "high precision"
viruses_all$best_coloring[viruses_all$toolcombo=="0 0 0 1 0 0"] <- "high precision"
viruses_all$best_coloring[viruses_all$toolcombo=="1 0 0 1 0 0"] <- "high precision"
viruses_all$best_coloring[viruses_all$toolcombo=="1 0 0 0 0 1"] <- "high precision"
# High MCC
viruses_all$best_coloring[viruses_all$toolcombo=="0 0 1 0 0 1"] <- "high MCC"
viruses_all$best_coloring[viruses_all$toolcombo=="0 1 1 0 0 1"] <- "high MCC"
viruses_all$best_coloring[viruses_all$toolcombo=="0 0 0 0 0 1"] <- "high MCC"
viruses_all$best_coloring[viruses_all$toolcombo=="1 0 1 1 0 1"] <- "high MCC"
viruses_all$best_coloring[viruses_all$toolcombo=="1 0 1 0 1 1"] <- "high MCC"
```

run commented-out lines after paper_figure_generation.Rmd if order changes

```{r}
# dput output from paper_figure_generation of toolcombo_vector created with the following commands
# toolcombo_MCC <- aggregate(accuracy_scores$MCC, list(accuracy_scores$toolcombo), mean)
# toolcombo_MCC <- toolcombo_MCC[order(toolcombo_MCC$x),]
# toolcombo_vector <- as.vector(toolcombo_MCC$Group.1)

# hard-coded
toolcombo_vector <- c("0 0 1 0 0 0", "0 1 0 0 0 0", "0 1 1 0 0 0", "0 1 0 0 1 0", 
"0 0 0 0 1 0", "0 0 1 0 1 0", "0 1 1 0 1 0", "0 1 0 1 1 0", "0 1 0 1 0 0", 
"0 0 0 1 1 0", "0 0 1 1 1 0", "0 1 1 1 1 0", "1 1 0 0 0 0", "0 1 1 1 0 0", 
"0 0 0 1 0 0", "0 0 1 1 0 0", "1 1 0 1 1 0", "1 1 0 1 0 0", "1 1 0 0 1 0", 
"1 1 0 1 1 1", "0 1 0 1 1 1", "1 1 0 0 1 1", "0 1 0 0 1 1", "1 1 0 1 0 1", 
"0 1 0 1 0 1", "1 1 0 0 0 1", "0 0 0 1 1 1", "1 1 1 0 0 0", "1 0 0 1 1 0", 
"1 0 0 1 1 1", "1 0 0 0 0 0", "0 0 0 0 1 1", "1 0 1 0 0 0", "1 0 0 0 1 0", 
"1 1 1 1 0 0", "1 0 0 0 1 1", "0 1 1 1 1 1", "1 1 1 1 1 0", "0 0 1 1 1 1", 
"0 1 0 0 0 1", "1 0 0 1 0 0", "0 0 0 1 0 1", "1 1 1 0 1 0", "1 0 1 1 1 0", 
"0 1 1 0 1 1", "1 0 1 1 0 0", "0 0 1 0 1 1", "0 1 1 1 0 1", "1 1 1 1 1 1", 
"1 0 0 1 0 1", "0 0 1 1 0 1", "1 0 1 0 1 0", "1 1 1 0 1 1", "1 0 0 0 0 1", 
"1 1 1 1 0 1", "1 0 1 1 1 1", "1 1 1 0 0 1", "1 0 1 0 1 1", "1 0 1 1 0 1", 
"0 0 0 0 0 1", "1 0 1 0 0 1", "0 1 1 0 0 1", "0 0 1 0 0 1")
```

```{r fig.align="center", fig.width = 5, fig.height = 2}
viruses_all$toolcombo <- factor(viruses_all$toolcombo, levels=toolcombo_vector)

tiff(filename='env_comparisons.tiff', units="in", width=15, height=6, res=300, compression = 'lzw')
ggplot(viruses_all, aes(x=toolcombo, y=prop_vir, 
                                  color=best_coloring, fill=best_coloring)) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=5, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Prop of Contigs >3kb Viral") +
  scale_fill_manual(name="",
                     values = alpha(rev(viridis(6)), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(rev(viridis(6)), 0.7)) +
  facet_wrap(~environment2)

dev.off()
```

```{r}
p.list = lapply(sort(unique(viruses_all$environment2)), function(i) {
ggplot(viruses_all[viruses_all$environment2==i,], aes(x=toolcombo, y=prop_vir, 
                                  color=best_coloring, fill=best_coloring)) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=5, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Prop of Contigs >3kb Viral") +
  scale_fill_manual(name="",
                     values = alpha(rev(viridis(6)), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(rev(viridis(6)), 0.7)) +
  facet_wrap(~environment2)
})
p.list
```

```{r}
viruses_all$type <- "metagenome"
viruses_all$type[viruses_all$environment2 %in% c("wastewater", "Lake_Erie_virome",
                                                 "Lake_Michigan")] <- "virome"
```

```{r}
ggplot(viruses_all[viruses_all$best_coloring %in% c("high MCC", "high recall"),], aes(x=environment2, y=prop_vir, 
                                  color=type, fill=type)) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Environment") +
  ylab("Proportion of Contigs >3kb Viral") +
  scale_fill_manual(name="",
                     values = alpha(magma(4)[2:3], 0.3)) +
  scale_color_manual(name="",
                     values = alpha(magma(4)[2:3], 0.7)) +
  facet_wrap(~best_coloring)
```

```{r}
imp_toolcombs <- viruses_all %>% group_by(best_coloring, environment2) %>%
  summarize(median=median(prop_vir),
            mean=mean(prop_vir),
            sd=sd(prop_vir)
            )
```

```{r}
ggplot(imp_toolcombs, aes(x=best_coloring, y=mean, 
                                  color=best_coloring, fill=best_coloring)) +
  geom_point(size=3) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
  ylab("Prop of Contigs >3kb Viral") +
  scale_fill_manual(name="",
                     values = alpha(rev(viridis(6)), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(rev(viridis(6)), 0.7)) +
  facet_wrap(~environment2)
```

```{r}
imp_toolcombs_wide <- viruses_all %>% group_by(best_coloring, environment) %>%
  summarize(median=median(prop_vir),
            mean=mean(prop_vir),
            sd=sd(prop_vir)
            ) %>%
  select(mean, environment, best_coloring) %>%
  pivot_wider(names_from=environment, values_from=mean)
```

# check how much tuning removal decreases viral ID by on average
```{r}
viruses_all <- viruses_all %>% 
  mutate(
    has_tnv = grepl('tnv', toolcombo2, fixed=TRUE),
    has_tv = grepl('tv', toolcombo2, fixed=TRUE),
  )
has_both <- rep(0, nrow(viruses_all))
has_both[viruses_all$has_tnv==TRUE] <- has_both[viruses_all$has_tnv==TRUE] + 0.5
has_both[viruses_all$has_tv==TRUE] <- has_both[viruses_all$has_tv==TRUE] + 0.5
viruses_all$has_both <- has_both

metagenomes <- viruses_all %>% subset(type == 'metagenome')
viromes <- viruses_all %>% subset(type == 'virome')

# how much does the tuning removal decrease the proportion identified as viral?
metaG_avg_decrease <- aggregate(metagenomes$prop_vir, list(metagenomes$has_tnv), mean)
virome_avg_decrease <- aggregate(viromes$prop_vir, list(viromes$has_tnv), mean)
metaG_avg_increase <- aggregate(metagenomes$prop_vir, list(metagenomes$has_tv), mean)
virome_avg_increase <- aggregate(viromes$prop_vir, list(viromes$has_tv), mean)

metaG_both <- aggregate(metagenomes$prop_vir, list(metagenomes$has_both), mean)
virome_both <- aggregate(viromes$prop_vir, list(viromes$has_both), mean)
```

# get total number of viruses and assemblies
```{r}
viruses_vs2_combo <- viruses_all[viruses_all$toolcombo == "1 0 1 0 0 1",]
viruses_vs2_combo$total_viruses <- viruses_vs2_combo$num_viruses / viruses_vs2_combo$prop_vir
```